---
title: "_Lab week 7 Map_"
author: "Lindsay Fisher"
date: "3/8/2022"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: TRUE
toc-title: "Table of Contents"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path= "../Output/")
```

# Load Libraries

I will be using the libraries: here, tidyverse, maps, mapdata, mapproj, and knitr.

Tidy Tuesday dataset used can be found [here.] (https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-11)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(maps)
library(mapdata)
library(mapproj)
library(knitr)
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")

```

# Meteorite Data

Below is a set of meteorite data from 2019-06-11 Tidy Tuesday on github.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
glimpse(meteorites)
world<-map_data("world") #have to create dataset called world from map data to call on it later
```

# Plot of Meteorite impacts

Below, the above data is plotted. Red points represent meteorite impacts and their geographical location. The plot only shows meteorite impacts since 1980.

```{r meteormap, echo=FALSE, warning=FALSE, message=FALSE}
metfilter<-meteorites %>% #creating new dataset
  filter(name_type == "Valid", #only takes vald data points
         fall == "Fell", #filters for meteorites that actually hit earth
         year >= 1980) #filters out all years before 1980
ggplot()+
    geom_polygon(data = world, #taking mapdata data - putting it in a map plot
               aes(x = long, #set x axis - actual column in dataset
                   y = lat, #set y axis
                   group = group, #makes the map appear with a clear outline
                   fill = region), #makes each country/region a different color
               color = "black")+ #outline is black
    guides(fill = FALSE)+ 
  theme(panel.background = element_rect(fill = "lightblue"))+ #changes map background to light blue
  labs( x = "Longitude", #changes labels to look cleaner
        y = "Latitude")+
 geom_point(data = metfilter, #layering a point plot with meteor data on top of map nd mapdata
             aes(x = lat,#sets x and y axes
                 y = long),
             color = "red", #makes points red - outside of aesthetics
             size=1)+ #sets size of points
    ggtitle("Global locations of meteorite impacts since 1980") #gives the plot a title


  


 


```

